---
title: "bbl_draft"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(lpSolve)
library(knitr)
library(zoo)
```

## Clean data

```{r}
# The web splits the data by batsmen, bowlers, wkp this creates duplicates
bbl <- read_csv('data/clean_bbl.csv')
bbl <- bbl[!duplicated(bbl[,1]),]
bbl
```

## Interpolate missing values

It seems that linear interpolation will work pretty well here so I'm just gonna use that.

```{r}
bbl <- bbl %>% arrange(desc(Cost))

cost <- bbl$Cost
avg <- ifelse(bbl$Avg_15_16 == 0, NA, bbl$Avg_15_16)
imp_loess <- predict(loess(avg ~ cost, span = 5))
imp_linear <- na.approx(cbind(cost, avg), rule = 2)[,2]

# cbind(cost, avg, imp_loess, imp_linear)

bbl$imputed_avg <- na.approx(cbind(cost, avg), rule = 2)[,2]
```

## Exploratory analysis

#### Cost vs Avg

```{r}
# Cost is a linear function of avg 2015-16 performance
ggplot(data = bbl) +
  geom_point(aes(Avg_15_16, Cost, colour = Position)) +
  theme_bw()
```

#### Cost vs total

```{r}
ggplot(data = bbl) +
  geom_point(aes(Total_15_16, Cost, colour = Position)) +
  theme_bw()
```

#### Cost vs imputed average

```{r}
ggplot(data = bbl) +
  geom_point(aes(imputed_avg, Cost, colour = Position)) +
  geom_label(aes(imputed_avg, Cost, fill = Team, label = Name)) +
  theme_bw()
```

## Linear programming

Maximise:

* Cost

Constraints:

* BAT == 7
* BWL == 7
* WKP == 2
* Cost <= 1,000,000
* Players == 16

```{r}
# instructions from http://pena.lt/y/2014/07/24/mathematically-optimising-fantasy-football-teams/

# create position dummies
bbl$bat <- as.numeric(grepl("BAT", bbl$Position))
bbl$bwl <- as.numeric(grepl("BWL", bbl$Position))
bbl$wkp <- as.numeric(grepl("WKP", bbl$Position))
bbl$allrnd <- as.numeric(grepl("BAT - BWL", bbl$Position))
bbl$plyr_count <- 1

# create constraints
nbat <- 7
nbwl <- 7
nwkp <- 2
nallrnd <- 2 # at least 2 all rounders
max_cost <- 1e6
nplayer <- 16 # must have a distinct number of players
min_games <- 40 # don't want too many players without good data


const_rhs <- c(nbat, nbwl, nwkp, nallrnd, max_cost, nplayer, min_games)
const_dir <- c(">=", ">=", "=", ">=", "<=", "=", ">=")
const_mat <- bbl %>% select(bat, bwl, wkp, allrnd, Cost, plyr_count, Games_15_16) %>% as.matrix %>% t

x <- lp("max", bbl$Cost, const_mat, const_dir, const_rhs, all.bin = TRUE, all.int = TRUE)
x
```


## Solve that g unit

```{r}
pick_team <- function(objective) {
  x <- lp("max", objective.in = objective, const_mat, const_dir, const_rhs, all.bin = TRUE, all.int = TRUE)
  bbl[as.logical(x$solution),]
}
```

### Select team based on maximising cost

```{r}
pick_team(bbl$Cost) %>% arrange(Position, desc(Cost)) %>% kable
```

### Select team based on maximising avg points in 15-16

```{r}
pick_team(bbl$Avg_15_16) %>% arrange(Position, desc(Cost)) %>% kable
```

### Select team based on maximising total points in 15-16

```{r}
pick_team(bbl$Total_15_16) %>% arrange(Position, desc(Cost)) %>% kable
```

### Select team based on maximising imputed average

```{r}
pick_team(bbl$imputed_avg) %>% arrange(Position, desc(Cost)) %>% kable
```

### No point having two good keepers

```{r}
# instructions from http://pena.lt/y/2014/07/24/mathematically-optimising-fantasy-football-teams/

# create position dummies
bbl$bat <- as.numeric(grepl("BAT", bbl$Position))
bbl$bwl <- as.numeric(grepl("BWL", bbl$Position))
bbl$wkp <- as.numeric(grepl("WKP", bbl$Position))
bbl$allrnd <- as.numeric(grepl("BAT - BWL", bbl$Position))
bbl$plyr_count <- 1

# create constraints
nbat <- 7
nbwl <- 7
nwkp <- 1
nallrnd <- 2 # at least 2 all rounders
max_cost <- 980000
nplayer <- 15 # must have a distinct number of players
min_games <- 50 # don't want too many players without good data


const_rhs <- c(nbat, nbwl, nwkp, nallrnd, max_cost, nplayer, min_games)
const_dir <- c(">=", ">=", "=", ">=", "<=", "=", ">=")
const_mat <- bbl %>% select(bat, bwl, wkp, allrnd, Cost, plyr_count, Games_15_16) %>% as.matrix %>% t

x <- lp("max", bbl$Cost, const_mat, const_dir, const_rhs, all.bin = TRUE, all.int = TRUE)
x
```


### Select team based on maximising cost

```{r}
pick_team(bbl$Cost) %>% arrange(Position, desc(Cost)) %>% kable
```

### Select team based on maximising avg points in 15-16

```{r}
pick_team(bbl$Avg_15_16) %>% arrange(Position, desc(Cost)) %>% kable
```

### Select team based on maximising total points in 15-16

```{r}
pick_team(bbl$Total_15_16) %>% arrange(Position, desc(Cost)) %>% kable
```

### Select team based on maximising imputed average

```{r}
pick_team(bbl$imputed_avg) %>% arrange(desc(Cost)) %>% kable
```